Windowed residual-minimization model reduction for dynamical systems

### 2019-09-19 main.tex
small progress

%\begin{itemize}
%\item Galerkin is inaccurate/unstable for non-symmetric static systems -- motivates residual minimization
%\item Residual minimization (stems from FEM): minimizes the residual of the full-order model over the range of the trial space -- originally developed for static systems
%\item Can guarantee stability for linear static systems 
%\item Extension to time-varying systems is challenging 
%\begin{itemize}
%\item What is the definition of the ODE residual?
%\item Over what time-interval is the residual being minmized?
%\end{itemize}
%\end{itemize} 
Least-squares residual minimization as a numerical solution technique for (partial) differential equations was pioneered in the finite element community and led to what is known as the least-squares finite element method (LSFEM)~\cite{ls_review,bochev_leastsquares}. Motivated by the fact that the Galerkin method can be inaccurate and unstable for nonlinear and non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic functional comprising the residual norm of the partial differential equation of interest and (2.) leveraging variational principles to derive the stationary conditions of the quadratic functional; i.e, LSFEM can be thought of as finding the solution within a finite element trial space the minimizes the $2$-norm of the PDE residual. Least-squares methods often exhibit numerous advantages (e.g., least-squares system are symmetric and positive definite) over the more standard Galerkin discretization.

Residual minimization methods were extended-to and developed-for reduced-order models by a variety of researchers. Similar to LSFEM, residual minimization approaches for reduced-order models comprise finding the solution within a low-dimensional trial space that minimizes the residual norm of the full-order model. Residual minimization approaches were first applied to static (i.e., no time-dependence) reduced-order models in~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis} and then extended to dynamical systems in~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}. Specifically, Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg} formulated the residual minimization problem for dynamical systems by sequentially minimizing the \textit{fully discrete} residual (i.e., the residual after temporal discretization) at each time-step instance. This formulation is most commonly known as the \textit{Least--Squares Petrov--Galerkin} (LSPG) method. Ref.~\cite{carlberg_lspg_v_galerkin} provided numerous analyses on this formulation and examined connections with the standard Galerkin method and Petrov--Galerkin methods. Critically, it was shown that (1.) for various time-stepping methods LSPG has an equivalent continuous form that corresponds to a Petrov--Galerkin projection and (2.) the standard Galerkin method minimizes the time-instantaneous residual of the full-order model. Numerous numerical experiments have demonstrated that LSPG yields more stable (although not always more accurate) solutions than the Galerkin method~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}. The common physical explanation for this improved performance is that, by minimizing the fully discrete residual, LSPG minimizes the residual over a \textit{finite-sized time-window}, rather than instantaneously as is done in the Galerkin method.

The LSPG approach only guarantees residual minimization over a single time-step, and as such is not associated with residual minimization of the full dynamical system.  
Indeed, LSPG is not equipped with \textit{a priori} guarantees of stability or convergence, even for linear time-invariant systems~\cite{bui_thesis}. Further, as it is defined at the fully discrete level, the performance of LSPG is subject to the time-step size: changing the time-step size changes the window over which the residual is minimized \textit{as well as} the temporal discretization error. As a result, LSPG provides optimal results at an intermediary time-step size. For too small a time-step, LSPG yields results similar to the Galerkin approach; in the limiting case of $\Delta t \rightarrow 0$, LSPG recovers Galerkin. For too-large a time-step, the accuracy of LSPG degrades. It is unclear if this is due to the numerical error associated with enlarging the time-step size, or rather if it is due to the size of the window the residual is being minimized over. The LSPG approach is additionally sensitive to the time discretization scheme. When used with explicit temporal discretization, for instance, LSPG again reverts to Galerkin. To gain the enhanced stability associated with LSPG, the underlying time-discretization scheme must be implicit with an appropriate time-step. This can increase the cost of the reduced-order model, especially in the case where the optimal time-step required by LSPG is small.

>>>>

%\begin{itemize}
%\item Galerkin is inaccurate/unstable for non-symmetric static systems -- motivates residual minimization
%\item Residual minimization (stems from FEM): minimizes the residual of the full-order model over the range of the trial space -- originally developed for static systems
%\item Can guarantee stability for linear static systems 
%\item Extension to time-varying systems is challenging 
%\begin{itemize}
%\item What is the definition of the ODE residual?
%\item Over what time-interval is the residual being minmized?
%\end{itemize}
%\end{itemize}

\KTC{Why are we reviewing LSFEM? This isn't a ROM. Don't think it's relevant}
Least-squares residual minimization as a numerical solution technique for
partial differential equations was pioneered in the finite-element community
and led to what is known as the least-squares finite-element method
(LSFEM)~\cite{ls_review,bochev_leastsquares}. 
\KTC{Specifically, LSFEM was designed to solve the inf--sup stability problem.
Be more precise.}
Motivated by the fact that the
Galerkin method can be inaccurate and unstable for nonlinear and
non-self-adjoint systems, the LSFEM involves (1.) forming a quadratic
functional comprising the residual norm of the partial differential equation
of interest and (2.) leveraging variational principles to derive the
stationary conditions of the quadratic functional; i.e, LSFEM can be thought
of as finding the solution within a finite element trial space the minimizes
the $2$-norm of the PDE residual. Least-squares methods often exhibit numerous
advantages (e.g., least-squares system are symmetric and positive definite)
\KTC{This is not an `advantage'. It is guaranteed to be inf--sup stable.}
over the more standard Galerkin discretization.

Researchers also developed residual-minimization methods for model reduction.
Similar to LSFEM, residual minimization in this context associates with
computing the solution within a low-dimensional trial space that minimizes the
$\ell^2$-norm of the full-order-model residual. \KTC{haven't defined
`full-order model'} Researchers have developed such residual-minimizing
model-reduction methods for both static systems (i.e., systems without
time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with Galerkin projection. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be derived
from a Petrov--Galerkin projection executed on the time-continuous
full-order-model residual, and (2) LSPG and Galekrin projection are equivalent
in the limit as the time step goes to zero (i.e., Galerkin projection
minimizes the time-instantaneous full-order model residual). Numerous
numerical experiments have demonstrated that LSPG projection often yields more
accurate and stable solutions than Galerkin
projection~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantanously), LSPG projection computes solutions that are more
accurate over a larger part the trajectory compared with Galerkin projection.

\KTC{HERE}
The LSPG approach only guarantees residual minimization over a single
time-step, and as such is not associated with residual minimization of the
full dynamical system.  Indeed, LSPG is not equipped with \textit{a priori}
guarantees of stability or convergence, even for linear time-invariant
systems~\cite{bui_thesis}. Further, as it is defined at the fully discrete
level, the performance of LSPG is subject to the time-step size: changing the
time-step size changes the window over which the residual is minimized
\textit{as well as} the temporal discretization error. As a result, LSPG
provides optimal results at an intermediary time-step size. For too small a
time-step, LSPG yields results similar to the Galerkin approach; in the
limiting case of $\Delta t \rightarrow 0$, LSPG recovers Galerkin. For
too-large a time-step, the accuracy of LSPG degrades. It is unclear if this is
due to the numerical error associated with enlarging the time-step size, or
rather if it is due to the size of the window the residual is being minimized
over. The LSPG approach is additionally sensitive to the time discretization
scheme. When used with explicit temporal discretization, for instance, LSPG
again reverts to Galerkin. To gain the enhanced stability associated with
LSPG, the underlying time-discretization scheme must be implicit with an
appropriate time-step. This can increase the cost of the reduced-order model,
especially in the case where the optimal time-step required by LSPG is small.
